Cheminformatics: The Computer Science of Chemical Discovery

### 2011-11-12 rebuttal.tex

\comment{P5, c2, l22: I would mention naive Bayesian classifiers as well as
this is perhaps the most widely applied method in the field.}
\response{We have included the Na\"{i}ve Bayes as well}

\comment{P5, c2, l42: formal citation for one of Hopfinger?s papers is
  required here.}
\response{The relevant reference has been added}

\comment{P5, c2, l58: this assertion is made with no evidence to back
  it up. Why should multi-target models be more reliable? It is not
  clear to me that they should.}\marginpar{\textbf{Needs more work}}
\response{We have rephrased this statement to be less dogmatic}

\comment{P6, c1, l7: I think it is important for the authors to cover ELNs by using actual recorded data. Do we know that many scientists changed to ELNs ?many years ago? as stated? Is there a reference for this or data to back it up?}
\response{We have expanded our original wording, and we now cite two references about the use and growth of ELNs in chemistry and elsewhere.}

\comment{P6, c1, l7: the authors do not mention the leading players in ELNs (e.g. CambridgeSoft, Accelrys, etc.). Perhaps this is because open-source ELNs do not exist, or are in their infancy? I think the key software houses that develop and supply these ELN systems should be mentioned. Furthermore, systems such as Reaxys should be mentioned to ensure that it is understood that these systems do exist to provide these results to these searches and are used daily by chemists.}
\response{We now mention CambridgeSoft, Accelrys and Reaxys in the article, and provide a citation to a review of 35 ELNs that was published in November 2011.}

### 2011-11-12 paper.tex
Added in Aarons edits

\begin{figure}
\centering
\includegraphics[height=2in]{chemstructs.png}
\caption{A hierarchy of chemical structure representations, increasing
  in information content from left to right and top to bottom. A
  chemists traditional 2D depiction (A), a labeled graph (B), a 3D
  conformation (C) and a surface represention (D). All show differing
  (and valid) aspects of the same molecule and are suited for
  differing purposes. The choice of representation is guided by the
  intended use of the structure data.}
\label{figure:chemical-structures}
\end{figure}

@@

The relevance of scientific knowledge management is increasing not
only for reducing the risk in current research, but also for enabling
new collaboration/innovation opportunities with internal partners and
a rapidly growing number of external/public partners. In fact, in
cheminformatics and chemistry, scientists already switched many years
ago from paper lab-notebooks to online collaboration platforms, called
ELNs (Electronic Lab Notebooks). So, what were the drivers behind
chemists, and companies, adopting ``Enterprise 2.0,'' a social online
collaboration culture?

@@

ELNs are typically much more accepted and used within the
cheminformatics domain than in the bioinformatics domain, which is
picking up speed lately, the question being why this is the case. In
chemistry, every experiment requires a lot of expert
knowledge. Therefore, approaches for reducing risks by communicating
with fellow chemists about similar reactions or compounds is
critical. This requires capturing the reaction details, compound
details, and time consuming analysis details which ensure quality
data, such that these data can be searched by others.

It is worth mentioning that chemistry is traditionally a field where many people
have been used to huge bookshelves of reactions and compounds. Chemists changed
to online ELN collaboration not only to increase efficiency, but also
because it allowed collaboration with trusted colleagues. Third party paper
catalogs and academic publications proved to be
inefficient, and even more critically, they did not take compounds of trusted
colleagues into account. Overall the change of management to ELNs, similar to
``Enterprise 2.0,'' required delivering on the promise to increase
interconnectivity, and to improve collaboration
opportunities, and indeed it already performed these functions many years ago.

>>>>

\begin{figure}
\centering
\includegraphics[height=2in]{chemical-structures.png}
\caption{A hierarchy of chemical structure representations, increasing
  in information content from left to right and top to bottom. A
  chemists traditional 2D depiction (A), a labeled graph (B), a 3D
  conformation (C) and a surface represention (D). All show differing
  (and valid) aspects of the same molecule and are suited for
  differing purposes. The choice of representation is guided by the
  intended use of the structure data.}
\label{figure:chemical-structures}
\end{figure}

@@

The relevance of scientific knowledge management is increasing not
only for reducing the risk in current research, but also for enabling
new collaboration/innovation opportunities with internal partners and
a rapidly growing number of external/public partners. In
cheminformatics and chemistry, scientists already switched many years
ago from paper lab-notebooks to online collaboration platforms, called
ELNs (Electronic Lab Notebooks). So, what were the drivers behind
chemists, and companies, adopting ``Enterprise 2.0,'' a social online
collaboration culture?

@@

While ELNs technically existed as long ago as the 1990s, chemists
adopted their use in large numbers in 2004 and 2005, and the market is
still growing~\cite{ELNstatus}. As one report notes, ``The initial
drive for the development of ELNs came from the field of chemistry,
perhaps driven by the early adoption by chemists of computer
technologies for drawing chemical structures and for storing and
searching them graphically within databases.... The pharmaceutical
industry shifted from a position of `if we get an ELN' to `when we get
an ELN'.''~\cite{ELNstatus} ELNs are typically much more accepted and
used within the cheminformatics domain than in the bioinformatics
domain, which is picking up speed lately, the question being why this
is the case. In chemistry, every experiment requires a lot of expert
knowledge. Therefore, approaches for reducing risks by communicating
with fellow chemists about similar reactions or compounds is
critical. This requires capturing the reaction details, compound
details, and time consuming analysis details which ensure quality
data, such that these data can be searched by others.

At this point, ``electronic lab notebooks have replaced the
traditional paper lab book across the pharmaceutical industry,'' but
are not yet common in academia, because of their high
cost~\cite{WavingGoodbye}.  It is worth mentioning that chemistry is
traditionally a field where many people have been used to huge
bookshelves of reactions and compounds. Chemists changed to online ELN
collaboration not only to increase efficiency, but also because it
allowed collaboration with trusted colleagues. Third party paper
catalogs and academic publications proved to be inefficient, and even
more critically, they did not take compounds of trusted colleagues
into account. Overall the change of management to ELNs, similar to
``Enterprise 2.0,'' required delivering on the promise to increase
interconnectivity, and to improve collaboration opportunities.  The
leading players in this area include CambridgeSoft's E-Notebook and
Accelrys' Symyx Notebook, but there are at least 35 different
companies producing ELNs as of February 2011~\cite{ELNreview}, for use
in chemistry, biology, quality control, and other research.  Another
knowledge management tool specifically for chemists is the web-based
service Reaxys, founded in 2009, whose company goal is to provide ``a
single, fully integrated chemical workflow solution.''
